---
title: "Data Rescue: Blanco 2011"
format:
  html:
    toc: true
    eval: false #default to not evaluating the chunks, override this the template is completed
date: last-modified
date-format: YYYY-MMMM
bibliography:
  - Blanco2011_Methods.bib 
  - Blanco2011.bib 
authors:
  - id: ktoddbrown 
    name:
      given: Katherine
      family: Todd-Brown
    orcid: 0000-0002-3109-8130
    affiliation:
      - ref: uf-ees
    role: #TODO define these roles and add reference
      - transcription: review #secondary, review
   #   - standardization: lead #review
    #  - curation: lead #review

affiliations:
  - id: uf-ees
    name: University of Florida
    department: Environmental Engineering Sciences
    city: Gainesville
    state: Florida
    country: USA
    address: "365 Weil Hall \nP.O. Box 116580"
    postal-code: 32611
    url: https://essie.ufl.edu/ees/
  - id: uf
    name: University of Florida
    city: Gainesville
    state: Florida
    country: USA
    postal-code: 32611
    url: https://www.ufl.edu/
    
timelog:
  - activity: 
      description: add general UF affilation
      who: ktoddbrown #github username
      #when you started and ended, can be just a year or down to the minute
      start: 2026-01-29 T1530
      #start: YYYY-MM-DD THH:MM
      #alternatively duration can be combined with a being/end
      duration: P05M
---


```{r}
#| label: setup
#| include: true
#| warning: false
#| message: false
#| echo: false
#| eval: true


library(tidyverse) # data processing
library(kableExtra) # pretty tables
library(bibtex) # reading in BibTex files to R

# data files for level 0
methods.file <- "Blanco2011_Methods.md"
table1.file <- 'Blanco2011_Table1.csv'
table2.file <- 'Blanco2011_Table2.csv'
table3.file <- 'Blanco2011_Table3.csv'
figure1.file <- 'Blanco2011_Figure1.csv'

# Bibliography files
primaryCitation.file <- 'Blanco2011.bib'
methodsCitation.file <- 'Blanco2011_Methods.bib'

# read function; you will need to create this function after you have your data
#... rescue reviewed and before you finish your data harmonization.
readsource.file <- '../../02_DataHarmonization/readBlanco2011.R'
if(file.exists(readsource.file)){
  source(readsource.file)
}

#semantic files for level 1
of_variable.file <- '../../semantics/SoilDRaHVocabulary_of_variable.csv'
is_type.file <- '../../semantics/SoilDRaHVocabulary_is_type.csv'

```


# Data Summary

> Blanco, J.A., Imbert, J.B. & Castillo, F.J. Thinning affects Pinus sylvestris needle decomposition rates and chemistry differently depending on site conditions. Biogeochemistry 106, 397–414 (2011). https://doi.org/10.1007/s10533-010-9518-2

For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/260.


@Blanco2011 examines how three thinning intensities affect *Pinus sylvestris* needle mass loss and chemistry over 5 years in Mediterranean and Continental forests. The data shows that thinning slows decomposition and extends nutrient immobilization, with the Mediterranean forest having faster initial decay but the Continental forest reaching a higher final decomposition limit.

## Tables and figures

<!--- Write a brief discription of all the tables and figures in this source. This should include full variable names when possible. It is perfectly fine to restate the captions here or copy other text that feels complete. Please keep this to a single sentence for each table/figure.--->

- Table 

  1) Table 1: Site characteristics (mean ± standard error)
  2) Table 2: Soil chemical and physical properties in the study sites
  3) Table 3: Regressions of changes in remaining organic mass versus changes in litter quality and climate variables
  4) Table 4: Initial chemical composition of senescent needles. All values in mg g^-1 (mean ± SE, n = 3)
  5) Table 5: Mean ± standard error of different variables at the time net release of N started

- Figures:

  1) Figure 1: Mass loss in decomposing needles during 5 years of study (mean ± standard error, n = 9)
  2) Figure 2: Nutrient content (N, P, K, Ca) in decomposing needles (mean ± standard error, n = 9).
  3) Figure 3: Nutrient (Mg) and organic compounds content (C, cellulose, lignin) in decomposing needles (mean ± standard error, n = 9).
  4) Figure 4: Conceptual model of N release in decomposing leaf litter emerging from our results.

## Fit for purpose: DeAD

The Decomposition Across Drylands (DeAD) data collection is a new data effort targeting dryland decomposition studies. The purpose of this data collection is to develop better litter decay models for dryland ecosystems.

### Main Variables of Interest

- Location: Evaluated across two environments in Navarre, northern Spain: Aspurz (Mediterranean forest: 42°48'50" N, 1°52'30" W) and Garde (Continental forest: 42°42'31" N, 1°8'40" W). This is found in Table 1.
- Soil Carbon Stock: Organic Carbon (Horizon A) is 67.8 mg g⁻¹ in Aspurz and 57.3 mg g⁻¹ in Garde, from Blanco 2004. This is found in Table 2. 
- Duration: The experiment spanned 5 years. Litter bags were deployed in November 2000, and final samples were retrieved in October 2005. This is found within the Materials and Methods section, subsection Preparation and placement of litter bags. 
- Mass Loss: Found explicitly in Figure 1.
- Solar Exposure: solar radiation measurements could not be found explicitly in this paper, but canopy opening effects through different thinning intensities (0%, 20%, 30%) are shown. These intensities are found in Figure 1.

### Other Relevant Variables
- Litter Species: *Pinus sylvestris* L. (Scots pine) needles.
- Microsite Data: Can be found in section "Materials and methods", subsection "Study sites and experimental design".
- Litter Chemical Composition: Initial concentrations of C, N, P, K, Ca, Mg, Cellulose, and Lignin across all thinning treatments are in Table 4. Also, in Figure 3.
- Litter Bag Materials and Methods: The litter bags measured 15 cm × 20 cm, made from fiberglass coated with polyethylene with a mesh size of 1.5 mm. Each bag was filled with 10 ± 0.001 g of air-dried needles. This is found in Materials and Methods section, subsection Preparation and placement of litter bags.
- Age of Ecosystem: 32 years old for Aspurz and 37 years old for Garde at the start of the study. Mentioned in Table 1.
- Land Management: Publicly owned managed forest stands subjected to thinning treatments implemented in August and November 1999. Mentioned in subsection: Study sites and experimental design.

## Files

These files are in the Blanco2011 data rescue.

- [Readme](README_Blanco2011.qmd)
  + This is the primary file that documents the transcriptions and decision made during data rescue.
- [Blanco2011.bib](Blanco2011.bib)
  + citation for article transcribed
- [Blanco2011_Methods.bib](Blanco2011_Methods.bib)
  + citations for the methods section of the article
- [Blanco2011_Methods.md](Blanco2011_Methods.md)
  + methods transcribed from primary article
- [Blanco2011_Table1.csv](Blanco2011_Table1.csv)
  + table 1 from primary article with coordinates of site, mean temperature, climate type, tree stand data before thinning, etc.
- [Blanco2011_Table2.csv](Blanco2011_Table2.csv)
  + table 2 from primary article with soil chemical and physical properties in the study sites
- [Blanco2011_Table4.csv](Blanco2011_Table4.csv)
  + table 4 from primary article with initial chemical composition of senescent needles.
- [Blanco2011_Table5.csv](Blanco2011_Table5.csv)
  + table 5 from primary article with mean ± standard error of different variables at the time net release of N started
- [Blanco2011_Figure1.csv](Blanco2011_Figure1.csv)
  + figure 1 from primary article with mass loss in decomposing needles during 5 years of study
- [Blanco2011_Figure3.csv](Blanco2011_Figure3.csv)
  + figure 3 from primary article with nutrient (Mg) and organic compounds content
- [temp/](temp/)
    + scratch folder that will not be archived on GitHub, it should include local copies of the data resource (likely a PDF article) and any spreadsheets or other tools you used for the transcriptions.

# Data Rescue working notes 

## Data Rescue Plan

+ Lead
  - [ ] Read the article or data resource and the discription of the target project
  - [ ] Create a fresh fork of the primary repository https://github.com/ktoddbrown/SoilDRaH/tree/main, it might be useful to rename it 'SoilDRaH_Blanco2011'
  - [ ] Create a copy of this template folder and rename it to Blanco2011
  - [ ] Fill out Data summary section (except for the Files), the Data Rescue Plan (here!), and update the author section of the yml to reflect who is working on the rescue
  - [ ] Create a GitHub issue from the Data Rescue template and get feedback from a reviewer on the project (Dr Todd-Brown or a grad student typically).
  - [ ] Add the reviewer to the author list in the yml
+ Do
  - [ ] Create a spreadsheet with TableN and export it to csv
  - [ ] Extract points from FigureM and export it to csv
  - [ ] Transcribe or copy the method section
  - [ ] Insert references to bib entries
  - [ ] reformat equations and mathematical symbols as [LaTex](https://en.wikibooks.org/wiki/LaTeX/Mathematics)
  - [ ] reformat text using [Markdown](https://www.markdownguide.org/cheat-sheet/)
  - [ ] Create a bib file with the citations from the methods section
  - [ ] Push to your local fork and get a second transcription pushed to your local fork by someone else on the project
+ Measure
  - [ ] cross check with second transcriber for tables and methods, and resolve any differences
  - [ ] compile or render this ReadMe with all chunks up through the Level 0 data rescue section with `eval: true`
  - [ ] push to your local fork and create a pull request to the main branch of the primary repository
  - [ ] address any revisions in your pull request review
  - [ ] celebrate!

## Table N

<!--- Common table modifications include: inserting an extra column to pivot sub-headers, filling in NA to identify blanks, and adding Latex or Markdown to denote formatting.---->

Table N was modified ....

```{r}
#| label: table1Check
#| eval: false #remove when you have the file transcribed


read_csv(file = tableN.file, #read in the specific file
         skip = 1, #skip the first line with the caption
         #read everything as a character, don't worry about formatting yet
         col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = tableN.file, #read in that file again but...
                           n_max = 1, #only read the first line
                           col_names = 'caption', #call it the caption
                           #read everything as a character
                           col_types = cols(.default = col_character()),
                           show_col_types = FALSE #do not return messages
                           )$caption #and take it out of the table
  )
```

And is crosschecked with the secondary transcription

```{bash}
#| label: compareTableN
#| eval: false

# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.

diff --strip-trailing-cr --suppress-common-lines -y temp/Blanco2011_TableN_ABC.md Blanco2011_TableN.md
```

## Methods

Any different to the cross check for the methods section will be output after this line.

```{bash}
#| label: compareMethods
#| eval: false
# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.
diff --strip-trailing-cr --suppress-common-lines -y temp/Blanco2011_Methods_ABC.md Blanco2011_Methods.md
```

### Methods  - Text

The final text transcription from the Methods section is rendered below.

```{r}
#| label: readMethods
#| output: asis
#| eval: true

# This revised section will append two ### to any existing headers here so that it renders nicely with the other headers in the document.

methodsLines <- readLines(methods.file) |>
  str_replace_all(pattern = '^#', replacement = '####')

cat(paste0(methodsLines, collapse = '\n'))

```

<!----
## Citation notes from Methods

Below are the citation notes from the methods section.
This generally doesn't read well in the final report so commenting it out here but it is very useful for tracking citations when you are doing the transcriptions.
Some citations are missing, those are noted here and how best guess were generated.
Think of this as a lab notebook documenting how you found each citation.

- [ ] short citation
  + copy from bib
  + manually entered or direct export
---->

# Level 0 Data rescue

This level 0 data read parses the transcribed files with minimal cleaning or transformations.

```{r}
#| label: readLevel0
#This chunk has two purposes. 
#...1) Check the formatting by reading in everything 
#...2) Create a list of everything to process later in level 1

data.lvl0.ls <- list(
  #Read in a list of all the bib files
  citation = list(
    #Citation for the article transcriptions are pulled from
    primary = read.bib(file = primaryCitation.file), 
    #Citations for all referenced articles
    methods = read.bib(file = methodsCitation.file)
  ),
  #Read in the text transcription of the article's methods section
  method = read_lines(file = methods.file),
  #Read in the results as tables or figure transcriptions. This includes
  #...the caption as well as the tables themselves
  data = list(
    Table1 = list(
      #Read the caption as a text string. Captions are the first cell on 
      #...the first row.
      caption = read_csv(file = tableN.file,
                         col_types = cols(.default = col_character()),
                         n_max = 1, col_names = FALSE)$X1[1],
      #Read in all the data, skipping the first row with the caption and read
      #...in the table as character. This element is a tibble (data.frame).
      primary = read_csv(file = tableN.file,
                         col_types = cols(.default = col_character()),
                         skip = 1)
    ) #End 'Table1' element
    #Repeat for all tables and figures...
   ) #end 'data' element
)

```

The code is also in `readsource.file` and returned with the `dataLevel = 'level0'` argument.
Below verifies that the data from the function matches what is coded above.

```{r}
#| label: checkReadFunction_lvl0

if(file.exists(readsource.file)){
  datafunction.lvl0.ls <- readBlanco2011(dataDir = '.',
                                         dataLevel = 'level0')
  
  if(isTRUE(all.equal(datafunction.lvl0.ls, data.lvl0.ls))){
    print('Read function matches code here.')
  }else{
    print('There is a mismatch in the data objects between the code here and the read function.')
  }
}else{
  print('The read function does not exist.')
}
```

# Level 1 data - CURATION PROJECT

This data is identified as a data source for the ...
Broad classes of information that we are looking for include:

<!--- Include description of where the data is located in the source.--->
- Location: 
- Soil carbon stock: 
- Soil type: 
- Land use: 

Information from the methods and other sections are pulled into tables with line by line source citations.
This information can be values, units, methods, or other useful notes and is manually extracted from the transcriptions loaded into the level 0 data above.

The final format of the data frame here will include the following 

  + grouping ids
  + `of_variable`
  + `is_type`
  + `with_entry`
  + `from_source`

The `of_variable` being used here are:

 + citation
 
 <!--- include other variables being used, this should match vocabulary in the SoilDRaHVocabulary resources --->

```{r}
#| label: createLvl1
#| eval: false

#### Method-section only data ####
#Create a table for the data in the method-section that is not directly 
#...associated with any table. Often this is study level information.

studyMeta <- tribble(~of_variable, ~is_type, ~with_entry, ~from_source,
                   'region', 'value', 'Honaunau Forest on the southwestern slopes of Mauna Loa, island of Hawaii', paste('Method ln5:', paste(data.lvl0.ls$method[5], collapse = ' ')))

#### Table/Figure N ####
# Process the primary data from Table/Figure M by maintaining data groups and 
#...standardizing the variable name so that we can link this to the study 
#...metadata and data in the other tables when needed. Link in the methods 
#...and other information from the method-section as needed.

Table1Primary <- data.lvl0.ls$data$Table1$primary |>
  #Add a row id preserving row grouping -- not unique
  mutate(row_id = paste0('R', 1:n())) |>
  #Transform the table to long data by row group and column names.
  #...This allows us to then create the cross links between the tables.
  pivot_longer(cols = -row_id,
               names_to = 'column_name', values_to = 'with_entry',
               values_drop_na = TRUE) |> #drop missing values
  #Link the variable names to the column names using a switch statement
  mutate(of_variable = case_when(
           #Soil PH is given as an example below
           column_name == 'SOC' ~ 'soil_organic_carbon')) |>
  #Flag the source of this data as Table 1, this table is merged later with
  #...other data so this keeps track of where the data came from
  mutate(from_source = 'Table 1')

## Create the metadata for table one from the header information and cited methods
Table1Meta <- Table1Primary |>
  select(column_name, of_variable) |>
  unique() |>
  #Grab everything between the parentheses as units and attribute the source as the column names.
  mutate(unit = str_extract(column_name, pattern = '(?<=\\().*(?=\\))'),
         from_source = 'Table 1 column names.') |>
  #If there aren't units then drop the row
  filter(!is.na(unit)) |>
  #Don't keep the column name now that you have the units, cross link via the variable
  select(-column_name) |>
  #Stack into the table the methods for each variable that we can find
  bind_rows(
  tribble(~of_variable, ~method, ~from_source,
            'soil_organic_carbon', paste0(data.lvl0.ls$method[72:73], collapse = ' '), 'Methods ln72-73') ) |>
  #Stack onto the tables the controlled vocabulary used
  bind_rows(
    tribble(~of_variable, ~control_vocabulary, ~from_source,
            'soil_class', 'Histosol:USDA classification for histosol soil type|Andisols:USDA classification for andisol soil type', 'expert informed')
  ) |>
  # Push the dimensions of the variable into a single column. Making this a long table.
  pivot_longer(cols = c(unit, method, control_vocabulary),
               names_to = 'is_type',
               values_drop_na = TRUE,
               values_to = 'with_entry')

#### Create level 1 #####
## Pull everything together into by stacking the meta and primary data tables
data.lvl1.ls <- list(
  meta = bind_rows(studyMeta, Table1Meta),
  primary = bind_rows(Table1Primary)|>
    arrange(row_id, column_name,
            of_variable, is_type, with_entry, from_source)
  )

```

## Check for new variables

`of_variable` in the data files should be defined in the SoilDRaH Vocabulary files.

```{r}
#| label: check_of_variable

#Read in the list of current 'of_variable' entries
SoilDRaH_variable_list <- read_csv(file =of_variable.file,
                                   show_col_types = FALSE)

#compare the variables in the level 1 tables to the SoilDRaH variables
new_variables <- lapply(data.lvl1.ls, function(xx){
  setdiff(unique(xx$of_variable), 
          unique(SoilDRaH_variable_list$of_variable))
}) |>
  unlist()

#Check that there are no new variables
if(length(new_variables) == 0){
  print('All of_variables are in the SoilDRaH variable list. No additions required')
}else{
  cat(paste('Please start a issue for new variables [',
               new_variables, ']', sep = '', collapse = '\n'))
}
```

## Check for is_type

`is_type` in the data files should be defined in the SoilDRaH Vocabulary files.

```{r}
#| label: check_is_type

#read in the current types and what variable they are associated with
SoilDRaH_type_list <- read_csv(file = is_type.file,
                                   show_col_types = FALSE) |>
  #remove the project definition and only keep the variable and type
  select(of_variable, is_type) |>
  #variables are pushed together in the type defined table. Separate the
  #...entries here
  separate_longer_delim(cols = 'of_variable', delim = ',')

#Pull all the of_variable-is_type pairs from the source data
source_pairs <- tibble()

#If there are a large number of entries, consider replacing this for-loop.
#...However this avoids having to install the plyr or other libraries, and 
#...there are generally no more then 3-5 tables
for(dataTable in data.lvl1.ls){
  source_pairs <- dataTable |>
    select(of_variable, is_type) |> 
    unique() |>
    bind_rows(source_pairs)
}

undefined_data_type <- source_pairs |>
  #subset for variables that are not 'incomplete'
  semi_join(SoilDRaH_variable_list |>
              filter(project_status != 'incomplete'),
            by = join_by(of_variable)) |>
  #check for of_variable-is_type tuples that are not in the SoilDRaH type definitions
  anti_join(SoilDRaH_type_list,
            by = join_by(of_variable, is_type))

#Tell the user what you found.
if(nrow(undefined_data_type) == 0){
  print('All of_variables-is_types are in the SoilDRaH type list. No additions required')
}else{
  print(paste0('Please start a issue for new of_variable-is_type [', paste(paste(undefined_data_type$of_variable, undefined_data_type$is_type, sep = '-'), collapse = ','), ']'))
}
```

## Check function match

Check for an exact match with the read function.

```{r}
#| label: checkReadFunction_lvl1

if(file.exists(readsource.file)){
  datafunction.lvl1.ls <- readBlanco2011(dataDir = '.',
                                         dataLevel = 'level1')
  
  if(isTRUE(all.equal(datafunction.lvl0.ls, data.lvl0.ls))){
    print('Read function output matches ReadMe code for level 1.')
  }else{
    print('There is a mismatch in the level 1 data objects between the code here and the read function.')
  }
}else{
  print('The read function does not exist.')
}

```


# References

<!---- This space will be filled with the citations when rendered. Leave this blank. ---->

