---
title: "Data Rescue: Idol 2007"
format:
  html:
    toc: true
date: last-modified
date-format: YYYY-MMMM
bibliography:
  - Idol2007_Methods.bib
  - Idol2007.bib 
authors:
  - id: ktoddbrown 
    name:
      given: Katherine
      family: Todd-Brown
    orcid: 0000-0002-3109-8130
    affiliation:
      - ref: uf-ees
    role: 
      - transcription: review 
      - standardization: review
    #  - curation: lead #review
  - id: waldera 
    name:
      given: Ashlynn
      family: Walder
    affiliation:
      - ref: uf-ees
    role: 
      - transcription: lead 
  - id: SavaScott 
    name:
      given: Savannah
      family: Scott
    affiliation:
      - ref: uf-ees
    role: 
      - transcription: secondary 
      - standardization: lead
    #  - curation: lead #review

affiliations:
  - id: uf-ees
    name: University of Florida
    department: Environmental Engineering Sciences
    city: Gainesville
    state: Florida
    country: USA
    address: "365 Weil Hall \nP.O. Box 116580"
    postal-code: 32611
    url: https://essie.ufl.edu/ees/
  - id: uf
    name: University of Florida
    city: Gainesville
    state: Florida
    country: USA
    postal-code: 32611
    url: https://www.ufl.edu/
    
timelog:
  - activity: 
      description: Primary Transcription
      who: waldera
      start: 2025-04-01
      end: 2025-05-24
      duration: PT3H
  - activity: 
      description: Secondary Transcription and PR
      who: SavaScott
      start: 2025-04-01
      end: 2025-05-24
      duration: PT5H
---

```{r}
#| label: setup
#| include: true
#| warning: false
#| message: false
#| echo: false
#| eval: true

library(tidyverse) # data processing
library(kableExtra) # pretty tables
library(bibtex) # reading in BibTex files to R

# data files for level 0
methods.file <- "Idol2007_Methods.md"
table1.file <- 'Idol2007_Table1.csv'
table2.file <- 'Idol2007_Table2.csv'
table3.file <- 'Idol2007_Table3.csv'

# Bibliography files
primaryCitation.file <- 'Idol2007.bib'
methodsCitation.file <- 'Idol2007_Methods.bib'

# read function; you will need to create this function after you have your data
#... rescue reviewed and before you finish your data harmonization.
readsource.file <- '../../02_DataHarmonization/readIdol2007.R'
if(file.exists(readsource.file)){
  source(readsource.file)
}

#semantic files for level 1
of_variable.file <- '../../semantics/SoilDRaHVocabulary_of_variable.csv'
is_type.file <- '../../semantics/SoilDRaHVocabulary_is_type.csv'

```

# Data Summary

> Idol, T., Baker, P. J., & Meason, D. (2007). Indicators of forest ecosystem productivity and nutrient status across precipitation and temperature gradients in Hawaii. Journal of Tropical Ecology, 23, 693 - 704. https://doi.org/10.1017/s0266467407004439

For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/83.

@Idol2007 examines how precipitation and temperature affect forest productivity, soil nutrients, and leaf nutrients in Hawaiian *Acacia koa* forests. It explores the relationships between climate, nutrient availability, and forest structure across independent elevation and rainfall gradients.

## Tables and figures

- Table 1: Site characteristics for Acacia koa stands on the Island of Hawaii
- Table 2: Basal area and soil and leaf nutrients across a Mean Annual Precipitation (MAP) gradient for Acacia koa stands at 1600–1650 m elevation
- Table 3: Stand productivity and soil and leaf nutrients from 1200–2050 m asl at Hawaii Volcanoes National Park (HAVO)

- Figure 1: Location of Acacia koa study plots on the Island of Hawaii
- Figure 2: Influence of leaf mass per area and soil C on measurements of leaf and soil N
- Figure 3: Influence of soil P on stand basal area and soil organic matter


## Fit for purpose: HiCSC

This data is identified as a data source for the Hawai'i Soil Organic Carbon Database https://osf.io/hmtv6/ as part of the HiCSC.
The [Hawaiʻi Climate Smart Partnership](https://climatesmarthawaii.org/) seeks to accelerate the implementation of climate-smart practices centered on Hawaiʻi-based producers and ancestral practitioners who are supported by advocacy, incentives, technical assistance, and market development.
As part of this project, we need to build a data base of soil carbon stocks found on different Hawaiian soil types under specific land uses and practices.


- Location: Keauhou Ranch (KR) 1600–1800 m elevation, Hawaii Volcanoes National Park (HAVO) 1200–2050 m elevation, and Honomalino (HONO) 1600 m elevation
- Soil carbon stock: Soil carbon stock is not explicitly provided, but total soil carbon content and SOM was determined from oven-dried samples in a Carlo Erba NA 2500 elemental analyser and the results are displayed in Table 2.
- Soil type: Andisols or as Histosols with andic properties
- Land use: Detailed land use in the Materials and Methods section.

# Files

These files are in the @Idol2007 data rescue.

- [Readme](README_Idol2007.qmd)
  + This is the primary file that documents the transcriptions and decision made during data rescue.
- [Idol2007.bib](Idol2007.bib)
  + citation for article transcribed
- [Idol2007_Methods.bib](Idol2007_Methods.bib)
  + citations for the methods section of the article
- [Idol2007_Methods.md](Idol2007_Methods.md)
  + methods transcribed from primary article
- [Idol2007_Table1.csv](Idol2007_Table1.csv)
  + table 1 from primary article with the site characteristics for *Acacia koa* stands on the Island of Hawaii. 
- [Idol2007_Table2.csv](Idol2007_Table2.csv)
  + table 2 from primary article with the basal area and soil and leaf nutrients across a Mean Annual Precipitation (MAP) gradient for *Acacia koa* stands at 1600–1650 m elevation.
- [Idol2007_Table3.csv](Idol2007_Table3.csv)
  + table 3 from primary article with the stand productivity and soil and leaf nutrients from 1200–2050 m asl at Hawaii Volcanoes National Park (HAVO).

# Data Rescue working notes

## Data Rescue Plan

+ Lead
  - [x] Read the article or data resource and the discription of the target project
  - [x] Create a fresh fork of the primary repository https://github.com/ktoddbrown/SoilDRaH/tree/main, it might be useful to rename it 'SoilDRaH_Idol2007'
  - [x] Create a copy of this template folder and rename it to Idol2007
  - [x] Fill out Data summary section (except for the Files), the Data Rescue Plan (here!), and update the author section of the yml to reflect who is working on the rescue
  - [x] Create a GitHub issue from the Data Rescue template and get feedback from a reviewer on the project (Dr Todd-Brown or a grad student typically).
  - [x] Add the reviewer to the author list in the yml
+ Do
  - [x] Create a spreadsheet with TableN and export it to csv
  - [x] Extract points from FigureM and export it to csv
  - [x] Transcribe or copy the method section
  - [x] Insert references to bib entries
  - [x] reformat equations and mathematical symbols as [LaTex](https://en.wikibooks.org/wiki/LaTeX/Mathematics)
  - [x] reformat text using [Markdown](https://www.markdownguide.org/cheat-sheet/)
  - [x] Create a bib file with the citations from the methods section
  - [x] Push to your local fork and get a second transcription pushed to your local fork by someone else on the project
+ Measure
  - [x] cross check with second transcriber for tables and methods, and resolve any differences
  - [x] compile or render this ReadMe with all chunks up through the Level 0 data rescue section with `eval: true`
  - [x] push to your local fork and create a pull request to the main branch of the primary repository
  - [x] address any revisions in your pull request review
  - [x] celebrate!


## Table 1

Table 1 was modified to include the footer in the header

```{r}
#| label: table1Check


read_csv(file = table1.file,
         skip = 1, 
        
         col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = table1.file,
                           n_max = 1, 
                           col_names = 'caption',
                          
                           col_types = cols(.default = col_character()),
                           show_col_types = FALSE
                           )$caption 
  )
```

And is crosschecked with the secondary transcription

```{bash}
#| label: compareTable1
#| eval: false

# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.

diff --strip-trailing-cr --suppress-common-lines -y temp/Idol2007_Table1_SRS.md Idol2007_Table1.md
```

## Table 2

Table 2 was not modified during the transcription.

```{r}
#| label: table2Check


read_csv(file = table2.file,
         skip = 1, 
        
         col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = table2.file,
                           n_max = 1, 
                           col_names = 'caption',
                          
                           col_types = cols(.default = col_character()),
                           show_col_types = FALSE
                           )$caption 
  )
```

And is crosschecked with the secondary transcription

```{bash}
#| label: compareTable2
#| eval: false

# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.

diff --strip-trailing-cr --suppress-common-lines -y temp/Idol2007_Table2_SRS.md Idol2007_Table2.md
```


## Table 3

Table 3 was modified to include +- as ±. 

```{r}
#| label: table3Check


read_csv(file = table3.file,
         skip = 1, 
        
         col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = table3.file,
                           n_max = 1, 
                           col_names = 'caption',
                          
                           col_types = cols(.default = col_character()),
                           show_col_types = FALSE
                           )$caption 
  )
```

And is crosschecked with the secondary transcription

```{bash}
#| label: compareTable3
#| eval: false

# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.

diff --strip-trailing-cr --suppress-common-lines -y temp/Idol2007_Table3_SRS.md Idol2007_Table3.md
```

## Methods

Any different to the cross check for the methods section will be output after this line.

```{bash}
#| label: compareMethods
#| eval: false

diff --strip-trailing-cr --suppress-common-lines -y temp/Idol2007_Methods_SRS.md Idol2007_Methods.md
```

### Methods  - Text

The final text transcription from the Methods section is rendered below.

```{r}
#| label: readMethods
#| output: asis
#| eval: true

# This revised section will append two ### to any existing headers here so that it renders nicely with the other headers in the document.

methodsLines <- readLines(methods.file) |>
  str_replace_all(pattern = '^#', replacement = '####')

cat(paste0(methodsLines, collapse = '\n'))

```


<!----
## Citation notes from Methods

Below are the citation notes from the methods section.
Some citations are missing, those are noted here and how best guess were generated.

- [ ] short citation
  + copy from bib
  + manually entered or direct export

- [x] TOSI, J. A., WATSON, V. & BOLANOS, R. 2002. *Life zone maps: Hawaii Guam, American Samoa, Northern Mariana Islands, Palau, and the Federated States of Micronesia.* Joint Venture of the Tropical Science Center, San Jose, Costa Rica and the Institute of Pacific Islands Forestry, USDA Forest Service, Hilo, 
  + copy from bib
- [x] LOCKWOOD, J. P., LIPMAN, P. W., PETERSON, L. D. & WARSHAUER, R. F. 1988. *Generalized ages of surface lava flows of Mauna Loa Volcano, Hawaii.* U.S. Geological Survey Miscellaneous Publications, Map I- 1908, U.S. Government Printing Office, Washington, DC. 
  + copy from bib
- [x] GIAMBELLUCA, T. W., NULLET, M. A. & SCHROEDER, T. A. 1986. *Rainfall atlas of Hawaii.* Department of Land and Natural Resources, Division of Water and Land Development, Honolulu. 267 pp.
  + copy from bib
- [x] ASSMANN, E. 19 70. *The principles offorest yield studies.* Pergamon Oxford. 506 pp.
  + copy from bib
- [x] BAKER, P. J. & SCOWCROFT, P. G. 2005. Stocking guidelines for endemic Hawaiian hardwood, *Acacia koa*. *Journal of Tropical Forest Science* 17:610-624.
  + copy from bib
- [x] FRALISH, J. S. 1994. The effect of site environment on forest productivity in the Illinois Shawnee Hills. *Ecological Applications* 4:134-143.
  + copy from bib
- [x] STERBA, H. & MONSERUD, R. A. 1993. The maximum density concept applied to uneven-aged mixed-species stands. *Forest Science* 39:432-452.
  + copy from bib
- [x] COOPERBAND, L. R. & LOGAN, T. J. 1994. Measuring in situ changes in labile soil phosphorus with anion-exchange membranes. *Soil Society of America Journal* 58: 105-114.
  + copy from bib
- [x] DROHAN, P. J., MERKLER, D. J. & BUCK, B. J. 2005. Suitability plant root simulator probe for use in the Mojave Desert. *Soil Society of America Journal* 69: 1482-1491.
  + copy from bib
- [x] MCGRATH, D. A., COMERFORD, N. B. & DURYEA, M. L. 2000. Litter dynamics and monthly fluctuations in soil phosphorus availability in an Amazonian agroforest. *Forest Ecology and Management* 131:167-
181.
  + copy from bib
- [x] MYERS, R. G., SHARPLEY, A. N., THIEN, S. J. & PIERZYNSKI, G. M. 2005. Ion-sink phosphorus extraction methods applied on 24 soils from the continental USA. *Soil Science Society of America Journal* 69:511-521.
  + copy from bib
- [x] QIAN, P. & SCHOENAU, J. J. 2005. Practical applications of ion exchange resins in agricultural and environmental soil research. *Canadian Journal of Soil Science* 82:9-21.
  + copy from bib
- [x] JOHNSON, D. W., VERBURG, P. S. J. & ARNONE, J. A. 2005. Soil extraction, ion exchange resin, and ion exchange membrane measures of soil mineral nitrogen during incubation of a tallgrass prairie soil. *Soil Science Society of America Journal* 69:260-265.
  + copy from bib
- [x] HERNANDEZ-MORENO, J. M. & NEGRIN, M. A. 1998. Phosphorus extraction by anion exchange membranes (AEM) in Andisols. *Communications in Soil Science and Plant Analysis* 29:1771-1778.
  + copy from bib
- [x] VITOUSEK, P. M., MATSON, P. A. & TURNER, D. R. 1988. Elevational and age gradients in Hawaiian montane rainforest: leaf and soil nutrients. *Oecologia* 77:565-570.
  + copy from bib
- [x] PEARSON, H. L. & VITOUSEK, P. M. 2001. Stand dynamics, nitrogen accumulation, and symbiotic nitrogen fixation in regenerating stands of *Acacia koa. Ecological Applications* 11:1381-1394.
  + copy from bib
- [x] HEDLEY, M. J., STEWART, J. W. B. & CHAUHAN, B. S. 1982. Changes in inorganic and organic phosphorus soil fractions  induced by cultivation practices and by laboratory incubations. *Soil Science Society of America Journal* 46:970-976.
  + copy from bib
- [x] TIESSEN, H. & MOIR, J. 0. 1993. Characterization of available P by sequential extraction. Pp. 75-86 in Carter M. R. (ed.). *Soil sampling and methods of analysis.* CRC Press, Boca Raton
  + copy from bib
- [x] OLSEN, S. R. & SOMMERS, L. E. 1982. Phosphorus. Pp. 403-430 in Page A. L., Miller R. H. & Keeney D. R. (eds.). *Methods of soil analysis.* Part 2.
*Chemical and microbiological methods.* American Society of Agronomy, Madison.
  + copy from bib
- [x] MURPHY, J. & RILEY, J. P. 1962. A modified single solution method for the determination of phosphate in natural waters. *Analytica Chimica Acta* 27:31-36
  + copy from bib
- [x] HURLBERT, S. H. 1984. Pseudoreplication and the design of field experiments. *Ecological Monographs* 54:187-211.
  + copy from bib
---->

# Level 0 data

```{r readLevel0}

#| label: readLevel0
#This chunk has two purposes. 
#...1) Check the formatting by reading in everything 
#...2) Create a list of everything to process later in level 1

data.lvl0.ls <- list(
  citation = list(
    primary = read.bib(file = primaryCitation.file), 
    methods = read.bib(file = methodsCitation.file)
  ),
  method = read_lines(file = methods.file),
  data = list(
    Table1 = list(
      caption = read_csv(file = table1.file,
                         col_types = cols(.default = col_character()),
                         n_max = 1, col_names = FALSE)$X1[1],
      primary = read_csv(file = table1.file,
                         col_types = cols(.default = col_character()),
                         skip = 1)
    ),
    Table2 = list(
      caption = read_csv(file = table2.file,
                         col_types = cols(.default = col_character()),
                         n_max = 1, col_names = FALSE)$X1[1],
      primary = read_csv(file = table2.file,
                         col_types = cols(.default = col_character()),
                         skip = 1)
    ),
    Table3 = list(
      caption = read_csv(file = table3.file,
                         col_types = cols(.default = col_character()),
                         n_max = 1, col_names = FALSE)$X1[1],
      primary = read_csv(file = table3.file,
                         col_types = cols(.default = col_character()),
                         skip = 1)
    )
  )
)
```

The code is also in `readsource.file` and returned with the `dataLevel = 'level0'` argument.
Below verifies that the data from the function matches what is coded above.

```{r}
#| label: checkReadFunction_lvl0

if(file.exists(readsource.file)){
  datafunction.lvl0.ls <- readAuthorYYYY(dataDir = '.',
                                         dataLevel = 'level0')
  
  if(isTRUE(all.equal(datafunction.lvl0.ls, data.lvl0.ls))){
    print('Read function matches code here.')
  }else{
    print('There is a mismatch in the data objects between the code here and the read function.')
  }
}else{
  print('The read function does not exist.')
}
```

# Level 1 data - CURATION HiCSC

This data is identified as a data source for the HICSC.
Broad classes of information that we are looking for include:

- Location:  Keauhou Ranch (KR) 1600–1800 m elevation, Hawaii Volcanoes National Park (HAVO) 1200–2050 m elevation, and Honomalino (HONO) 1600 m elevation
- Soil carbon stock: Soil carbon stock is not explicitly provided, but total soil carbon content and SOM was determined from oven-dried samples in a Carlo Erba NA 2500 elemental analyser and the results are displayed in Table 2.
- Soil type: Andisols or as Histosols with andic properties
- Land use: Detailed land use in the Materials and Methods section.

Information from the methods and other sections are pulled into tables with line by line source citations.
This information can be values, units, methods, or other useful notes and is manually extracted from the transcriptions loaded into the level 0 data above.

The final format of the data frame here will include the following 

  + grouping ids
  + `of_variable`
  + `is_type`
  + `with_entry`
  + `from_source`

The `of_variable` being used here are:

 + citation
 + region
 + mean_annual_precipiation
 + land_use
 + soil_class
 + stand_age
 + site_name
 + elevation
 + mean_annual_temperature
 + stand_age
 + soil_age
 + stem_density

```{r}
#| label: createLvl1
#| eval: false

#### Method-section only data ####

studyMeta <- tribble(~of_variable, ~is_type, ~with_entry, ~from_source,
                   'region', 'value', 'slopes of the Mauna Loa Volcano', paste('Method ln9:', paste(data.lvl0.ls$method[9], collapse = ' ')),
                   'land_use', 'value', 'stands that had regenerated into almost pure A. *koa* forests after livestock exclusion and ungulate removal', paste('Method ln8:', paste(data.lvl0.ls$method[8], collapse = ' ')),
                   'soil_class', 'value','Andisols or as Histosols with andic properties reflective of the volcanic parent material',  paste('Method ln28:', paste(data.lvl0.ls$method[28], collapse = ' '),
                   'citation', 'value', format(data.lvl0.df$citation$primary), 'journal citation',
                   'doi', 'value', data.lvl0.df$citation$primary$doi, 'journal  citation'))
  

#### Table 1 ####

Table1Primary <- data.lvl0.ls$data$Table1$primary |>
  mutate(row_id = paste0('R', 1:n())) |>
  pivot_longer(cols = -row_id,
               names_to = 'column_name', values_to = 'with_entry',
               values_drop_na = TRUE) |> 
  mutate(of_variable = case_when(
           column_name == 'Site' ~ 'site_name',
           column_name == 'Elevation (m)' ~ 'elevation',
           column_name == 'MAP(mm)' ~ 'mean_annual_precipiation',
           column_name == 'MAT(°C)' ~ 'mean_annual_temperature',
           column_name == 'Stand age (y)' ~ 'stand_age',
           column_name == 'Soil age (y)' ~ 'soil_age',
           column_name == 'Density (stems ha<sup>-1</sup>)' ~ 'stem_density'
         )) |>
  mutate(from_source = 'Table 1')

Table1Meta <- Table1Primary |>
  select(column_name, of_variable) |>
 unique() |>
  mutate(unit = str_extract(column_name, pattern = '(?<=\\().*(?=\\))'),
         species = case_when(str_detect(column_name, 'A. koa') ~ 'Acacia koa',
                             .default = NA_character_),
         from_source = 'Table 1 column names.') |>
  filter(!is.na(unit) | !is.na(species)) |>
  
  bind_rows(
    tribble(~of_variable, ~method, ~from_source,
            'site_name', paste0(data.lvl0.ls$method[c(10,12,14)], collapse = ' '), 'Methods ln10,12,14',
            'elevation', paste0(data.lvl0.ls$method[19:21], collapse = ' '), 'Methods ln19-21',
            'mean_annual_precipiation', paste0(data.lvl0.ls$method[c(9,30,32)], collapse = ' '), 'Methods ln9,30,32',
            'mean_annual_temperature', paste0(data.lvl0.ls$method[c(19:20,30,35:26)], collapse = ' '), 'Methods ln19-20,30,35-36',
            'stand_age', paste0(data.lvl0.ls$method[22], collapse = ' '), 'Methods ln22',
            'soil_age', paste0(data.lvl0.ls$method[26:27], collapse = ' '), 'Methods ln26-27',
            'stem_density', paste0(data.lvl0.ls$method[23:24], collapse = ' '), 'Methods ln23-24'
            ) ) |>
  
  
 bind_rows(
    tribble(~of_variable, ~control_vocabulary, ~from_source,
          'soil_class', 'Andisols: USDA classification for soil developing in volcanic ejecta| Histosols: USDA taxonomic class of  soils forming in organic soil materials', 'USDA',
          'stand_type', '*A. koa*: pure stands of *Acacia koa* Gray', 'Introduction ln6',) |>
left_join(Table1Primary |>
                  select(column_name, of_variable) |>
                  unique(),
                by = join_by(of_variable)))|>
  pivot_longer(cols = c(unit, species, method, control_vocabulary),
               names_to = 'is_type',
               values_drop_na = TRUE,
               values_to = 'with_entry')

#### Table 2 ####

Table2Primary <- data.lvl0.ls$data$Table2$primary |>
  mutate(row_id = paste0('R', 1:n())) |>
  pivot_longer(cols = -row_id,
               names_to = 'column_name', values_to = 'with_entry',
               values_drop_na = TRUE) |> 
  mutate(of_variable = case_when(
           column_name == 'Site' ~ 'site_name',
           column_name == 'Elevation (m)' ~ 'elevation',
           column_name == 'MAP(mm)' ~ 'mean_annual_precipiation',
           column_name == 'MAT(°C)' ~ 'mean_annual_temperature',
           column_name == 'Stand age (y)' ~ 'stand_age',
           column_name == 'Soil age (y)' ~ 'soil_age',
           column_name == 'Density (stems ha<sup>-1</sup>)' ~ 'stem_density'
         )) |>
  mutate(from_source = 'Table 1')

Table1Meta <- Table1Primary |>
  select(column_name, of_variable) |>
 unique() |>
  mutate(unit = str_extract(column_name, pattern = '(?<=\\().*(?=\\))'),
         species = case_when(str_detect(column_name, 'A. koa') ~ 'Acacia koa',
                             .default = NA_character_),
         from_source = 'Table 1 column names.') |>
  filter(!is.na(unit) | !is.na(species)) |>
  
  bind_rows(
    tribble(~of_variable, ~method, ~from_source,
            'site_name', paste0(data.lvl0.ls$method[c(10,12,14)], collapse = ' '), 'Methods ln10,12,14',
            'elevation', paste0(data.lvl0.ls$method[19:21], collapse = ' '), 'Methods ln19-21',
            'mean_annual_precipiation', paste0(data.lvl0.ls$method[c(9,30,32)], collapse = ' '), 'Methods ln9,30,32',
            'mean_annual_temperature', paste0(data.lvl0.ls$method[c(19:20,30,35:26)], collapse = ' '), 'Methods ln19-20,30,35-36',
            'stand_age', paste0(data.lvl0.ls$method[22], collapse = ' '), 'Methods ln22',
            'soil_age', paste0(data.lvl0.ls$method[26:27], collapse = ' '), 'Methods ln26-27',
            'stem_density', paste0(data.lvl0.ls$method[23:24], collapse = ' '), 'Methods ln23-24'
            ) ) |>
  
  
 bind_rows(
    tribble(~of_variable, ~control_vocabulary, ~from_source,
          'soil_class', 'Andisols: USDA classification for soil developing in volcanic ejecta| Histosols: USDA taxonomic class of  soils forming in organic soil materials', 'USDA',
          'stand_type', '*A. koa*: pure stands of *Acacia koa* Gray', 'Introduction ln6',) |>
left_join(Table1Primary |>
                  select(column_name, of_variable) |>
                  unique(),
                by = join_by(of_variable)))|>
  pivot_longer(cols = c(unit, species, method, control_vocabulary),
               names_to = 'is_type',
               values_drop_na = TRUE,
               values_to = 'with_entry')


```

